Title

### 2024-11-23 paper/paper.qmd
unneeded file deleted paper rendered

---
title: "Title"
subtitle: "Subtitle"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/ChenmingZ2000/Loblaws-NoNameBrand-Grocery-Price-Trend-Analysis.git](https://github.com/ChenmingZ2000/Loblaws-NoNameBrand-Grocery-Price-Trend-Analysis.git).
        We gratefully acknowledge Jacob Filipp’s groceries dataset used in this analysis. The data is available at: [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/).
date: today
date-format: long
abstract: "abstract"
format: pdf
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{paper/tabularray}
pdf-engine: lualatex
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

The product pricing data for the "No Name" brand sold by Loblaws, used in this analysis, is included in the dataset provided by [@hammer], sourced from [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/). This dataset compiles detailed product information, including vendors, product names, brands, categories, and unit prices. Additionally, it includes time-series pricing records for each product, capturing timestamps of price changes, current prices, unit prices, and quantities sold. These records are periodically updated on the Hammer platform to reflect the latest pricing trends. This dataset forms the foundation for analyzing price trends, identifying unit variations, and categorizing products into specific types for further analysis. Table @tbl-productdata previews the information for the target products included in this analysis, starting from June 2024.

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `readxl` [@readxl], `dplyr` [@dplyr], `arrow` [@arrow], `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

In the real world, pricing data for consumer goods is a dynamic reflection of market trends, regional preferences, and competitive strategies. The prices of products fluctuate based on supply chain disruptions, seasonal demand, and retail strategies, making it essential to track and analyze this information systematically. The Hammer dataset serves as a resource in understanding these trends, offering a detailed collection of product pricing data sourced from leading retailers like Loblaws, Walmart, and Metro. This dataset is embedded with complexities arising from the data collection process, competitive pricing strategies, and evolving consumer behavior.

@@

- `Size`: Product size is a determinant of pricing, as larger sizes generally cost more due to the increased quantity. However, pricing strategies such as bulk discounts or price-per-unit variations can create non-linear relationships between size and price, making size an important factor to analyze.
- `Type`: Product type (e.g., snacks, seasoning, dairy) reflects category-level differences in pricing strategies. Different types of products may have varying production costs, supply chain logistics, and market demand, all of which contribute to price fluctuations.
- `Date`: Time plays an important role in grocery pricing due to reasons like seasonal demand, inflation, supply chain disruptions, and promotions. Analyzing the impact of time allows us to capture temporal trends and shifts in pricing behavior, such as rising costs during inflationary periods or discounts during specific seasons.

@@

Define yiy_i as the current price of the product in the ii-th observation, assumed to follow a Gamma distribution with mean $\mu_i$ and shape parameter $\phi$. This model accounts for variations across different product sizes, types, and temporal trends, with random intercepts for each product ID.

The linear predictor for the mean ($\mu_i$) of the Gamma distribution is defined as follows:
$$
\log(\mu_i) = \alpha + \beta_1 \cdot \text{size}_ i + \beta_2 \cdot \text{type}_i + \beta_3 \cdot \text{date}_i + u_{\text{product_id}[i]} \\
$$

where:
$y_i$ is the observed current price for the ii-th product.
$\mu_i$ is the predicted mean price for the ii-th product, modeled on a log scale.
$\alpha \sim \text{Normal}(2, 5)$ is the global intercept representing the baseline price.
$\beta_1 \sim \text{Normal}(0, 5)$ captures the effect of product size on pricing (e.g., Small, Large, Extra Large).
$\beta_2 \sim \text{Normal}(0, 5)$ represents the effect of product type (e.g., Snacks, Flour/Rice, Cooking Oil) on pricing.
$\beta_3 \sim \text{Normal}(0, 5)$ accounts for the temporal trend in prices over time.

Random effects:

$u_{\text{state}} \sim \text{Normal}(0, \sigma_{\text{state}})$

$v_{\text{pollster}} \sim \text{Normal}(0, \sigma_{\text{pollster}})$

$\gamma_{\text{pollster}} \sim \text{Normal}(0, \sigma_{\mathrm{pollster\_slope}})$

@@

- $u_{\text{state}[i]}$ is the random intercept for each state, allowing for state-level variations in Trump support.
- $v_{\text{pollster}[i]}$ is the random intercept for each pollster, capturing pollster-specific effects.
- $\gamma_{\text{pollster}[i]}$ is the random slope associated with the time variable ($\mathrm{days\_since\_Biden\_Withdrawal}$) for each pollster.

@@

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "Title"
subtitle: "Subtitle"
author: 
  - Chenming Zhao
thanks: "Code and data are available at: [https://github.com/ChenmingZ2000/Loblaws-NoNameBrand-Grocery-Price-Trend-Analysis.git](https://github.com/ChenmingZ2000/Loblaws-NoNameBrand-Grocery-Price-Trend-Analysis.git).
        We gratefully acknowledge Jacob Filipp’s groceries dataset used in this analysis. The data is available at: [https://jacobfilipp.com/hammer/](https://jacobfilipp.com/hammer/)"
date: today
date-format: long
abstract: "abstract"
format: pdf
number-sections: true
bibliography: references.bib
#header-includes:
#  - \usepackage{tabularray/tabularray}
pdf-engine: lualatex
---

@@

@@

The product pricing data for the "No Name" brand sold by Loblaws, used in this analysis, is included in the dataset provided by [@hammer]. This dataset compiles detailed product information, including vendors, product names, brands, categories, and unit prices. Additionally, it includes time-series pricing records for each product, capturing timestamps of price changes, current prices, unit prices, and quantities sold. These records are periodically updated on the Hammer platform to reflect the latest pricing trends. This dataset forms the foundation for analyzing price trends, identifying unit variations, and categorizing products into specific types for further analysis. Table @tbl-productdata previews the information for the target products included in this analysis, starting from June 2024.

To simulate, test, analyze, and clean the polling data, the statistical programming language R was employed [@citeR]. Key libraries that supported the data analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr], `arrow` [@arrow], `knitr` [@knitr], `ggplot2` [@ggplot2], `kableExtra` [@kableExtra] and `patchwork` [@patch]. Additionally, the `rstanarm` [@rstanarm] library was utilized to implement multilevel modeling for prediction purposes. Also this analysis inherits the folder structure from Professor Rohan Alexander from University of Toronto [@rohan].

## Measurement
In the real world, pricing data for consumer goods is a dynamic reflection of market trends, regional preferences, and competitive strategies. The prices of products fluctuate based on supply chain disruptions, seasonal demand, and retail strategies, making it essential to track and analyze this information systematically. The Hammer dataset serves as a resource in understanding these trends, offering a detailed collection of product pricing data sourced from leading retailers like Loblaws, Walmart, and Metro. This dataset is embedded with complexities arising from the data collection process, competitive pricing strategies, and evolving consumer behavior.

@@

**Size**: Product size is a determinant of pricing, as larger sizes generally cost more due to the increased quantity. However, pricing strategies such as bulk discounts or price-per-unit variations can create non-linear relationships between size and price, making size an important factor to analyze.
**Type**: Product type (e.g., snacks, seasoning, dairy) reflects category-level differences in pricing strategies. Different types of products may have varying production costs, supply chain logistics, and market demand, all of which contribute to price fluctuations.
**Date**: Time plays an important role in grocery pricing due to reasons like seasonal demand, inflation, supply chain disruptions, and promotions. Analyzing the impact of time allows us to capture temporal trends and shifts in pricing behavior, such as rising costs during inflationary periods or discounts during specific seasons.

@@

Define $y_i$ as the current price of the product in the i-th observation, assumed to follow a Gamma distribution with mean $\mu_i$ and shape parameter $\phi$. This model accounts for variations across different product sizes, types, and temporal trends, with random intercepts for each product ID.

The linear predictor for the mean ($\mu_i$) of the Gamma distribution is defined as follows:

$log(\mu_i)$ = $\alpha$ + $\beta_1 \cdot \text{size}_ i$ + $\beta_2 \cdot \text{type}_i$ + $\beta_3 \cdot \text{date}_i$ + $u_{\text{productID}[i]}$

@@

- $y_i$ is the observed current price for the i-th product.

- $\mu_i$ is the predicted mean price for the i-th product, modeled on a log scale.

- $\alpha \sim \text{Normal}(2, 5)$ is the global intercept representing the baseline price.

- $\beta_1 \sim \text{Normal}(0, 5)$ captures the effect of product size on pricing (e.g., Small, Large, Extra Large).

- $\beta_2 \sim \text{Normal}(0, 5)$ represents the effect of product type (e.g., Snacks, Flour/Rice, Cooking Oil) on pricing.

- $\beta_3 \sim \text{Normal}(0, 5)$ accounts for the temporal trend in prices over time.

Random Effects: $u_{\text{productID}} \sim \text{Normal}(0, \sigma_{\text{productID}})$

$u_{\text{productID}[i]}$ is the random intercept for each product ID, capturing product-specific variations in pricing.

@@

Our results are summarized in @tbl-model.

```{r, message=FALSE}
#| echo: false
#| eval: false
#| warning: false
#| message: false

@@

first_model <- readRDS(file = here::here("models/price_model.rds"))

```

```{r, message=FALSE}
#| echo: false
#| eval: false
#| label: tbl-model
#| tbl-cap: Model summary table for model which predicts Loblaws "No Name" grocery price in 2024
#| warning: false
#| message: false

modelsummary::modelsummary(
  list("Loblaws 'No Name' Price Gamma model" = first_model),
  statistic = "mad",
  fmt = 2,
  coef_map = c("Sigma[product_id × (Intercept),(Intercept)]", "(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice")
)
```
